Dataflow-Guided Filtering for Efficient and Adjustable Run-Time Monitoring

### 2014-05-30 introduction.tex
Revise introduction.

% Programmable hardware monitors
These run-time monitors introduce performance, power, and energy overheads.
Implementing run-time monitoring in software using binary instrumentation or
other similar methods has been shown to introduce high overheads. For example,
dynamic information flow tracking (DIFT) implemented in software suffers a 3.6x
slowdown \cite{lift-micro06}. Implementing monitors in hardware greatly decreases
the performance impact by performing monitoring in parallel to a program's
execution. A dedicated hardware implementation of DIFT reduces performance
overheads to just 1.1\% \cite{dift-asplos04}. However, fixed hardware loses
the programmability of software, and cannot be updated or changed.

% Filtering
Recent studies have shown that filtering out monitoring checks on clean or null
metadata can be used to reduce the overheads of monitoring \cite{fade-hpca14}. For example, for
an array bounds check, checking non-pointer accesses is unnecessary. In this
paper, we present an architecture that combines a hardware-based dataflow
tracking engine with a software-based monitor in order to filter out monitoring operations 
for null metadata. By keeping a separate set of flags
to track null metadata, we are able to handle the common case of null
metadata without the need to read the full metadata values. By propagating these flags, we
are also able to handle propagation of null metadata efficiently in hardware.
This filtering can be used to greatly reduce the overheads of monitoring.

% Adjustable overhead
There is a limit to how effective filtering is at reducing monitoring overheads. 
Traditionally, these overheads have been an all-or-nothing cost. If the
designer finds that the overheads exceed their desired budget, then monitoring
cannot be enabled.  
Thus, we propose to perform partial monitoring to enable further reduction of overheads.
 This is achieved by dynamically dropping
monitoring operations when overheads exceed a specified overhead budget. We prevent false
positives from partial monitoring by using the dataflow engine to track these dropped
monitoring flows. We investigate different policies for deciding which events
to drop for partial monitoring and show that there is a trade-off between
closely matching the overhead budget and the monitoring coverage achieved.

Although full monitoring coverage is always desired, trading off coverage for
reduced overheads can be very useful. For example, this adjustable
monitoring enables a level of protection even for systems where the full
monitoring overheads are too high. This is especially true for energy- or
power-constrained systems or soft real-time systems where the monitoring
overhead should not exceed energy/power limits or real-time deadlines.
Additionally, adjustable monitoring can be used for debugging purposes
especially in the context of sampling-based or cooperative debugging techniques
which expect low coverage per run but use a large number of runs and users to
collect debugging information~\cite{liblit-pldi05, chilimbi-asplos04,
greathouse-cgo11}.

>>>>

% Programmable hardware monitors
These run-time monitors introduce performance, power, and energy overheads.
% Implementing run-time monitoring in software using binary instrumentation or
% other similar methods has been shown to introduce high overheads. For example,
% dynamic information flow tracking (DIFT) implemented in software suffers a 3.6x
% slowdown \cite{lift-micro06}. 
Implementing monitors in dedicated hardware has shown to have very small
performance impacts. For example, a dedicated hardware implementation of
dynamic information flow tracking (DIFT) shows performance
overheads of only 1.1\% \cite{dift-asplos04}. However, fixed hardware loses
the programmability of software, and cannot be updated or changed. It also
requires for monitors to be chosen and implemented at hardware design time.
Recent studies \cite{lba-isca08} have proposed using parallel cores in a
multi-core system to perform monitoring. This allows monitoring to be done
using parallel hardware while having the flexibility and programmability of
software. However, due to the multiple cycles it may take for software to
monitor an instruction, the overheads can easily increase execution time
several fold.

% Two optimizations
In this paper, we present a hardware architecture that uses dynamic dataflow
tracking in order to reduce the overheads of programmable parallel monitoring.
This dataflow tracking engine is used to filter out unnecessary monitoring and
enable partial monitoring.

% Filtering
Recent studies have shown that filtering out monitoring checks on clean or null
metadata can be used to reduce the overheads of monitoring \cite{fade-hpca14}.
For example, for
an array bounds check, checking non-pointer accesses is unnecessary. 
In this paper, we use a hardware-based dataflow tracking engine to track and
filter out these null metadata monitoring operations.
By using the dataflow engine to track a separate set of flags
indicating null metadata, we are able to handle the common case of null
metadata without the need to read the full metadata values.

Although full monitoring coverage is always desired, trading off coverage for
reduced overheads can be very useful. For example, partial
monitoring can enable a level of protection even for systems where the full
monitoring overheads are too high. This is especially true for energy- or
power-constrained systems or soft real-time systems where the monitoring
overhead should not exceed energy/power limits or real-time deadlines.
Additionally, partial monitoring can be used for debugging purposes
especially in the context of sampling-based or cooperative debugging techniques
which expect low coverage per run but use a large number of runs and users to
collect debugging information~\cite{liblit-pldi05, chilimbi-asplos04,
greathouse-cgo11}.

% % Adjustable overhead
% There is a limit to how effective filtering is at reducing monitoring overheads. 
% Traditionally, these overheads have been an all-or-nothing cost. If the
% designer finds that the overheads exceed their desired budget, then monitoring
% cannot be enabled.  
In this paper, we show how our dataflow-guided architecture is also able to
enable partial monitoring. This is achieved by dynamically dropping monitoring
operations when overheads exceed a specified overhead budget. We prevent false
positives from partial monitoring by using the dataflow engine to track these
dropped monitoring flows. With this architecture, we investigate different
policies for deciding which events to drop for partial monitoring. These
different policies show a trade-off between closely matching the overhead
budget and the monitoring coverage achieved.
